% ============================================================
%  One Shell to Fix Them All: A Minimal LLM-Driven Coding Agent
%  with Plan--Execute Decomposition for Automated Bug Repair
%
%  ACL proceedings format
% ============================================================
\documentclass[11pt]{article}

% ACL style (change "review" to "final" for camera-ready)
\usepackage[review]{acl}

% Standard ACL packages
\usepackage{times}
\usepackage{latexsym}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{inconsolata}
\usepackage{graphicx}

% Additional packages for math, algorithms, and code listings
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{listings}
\usepackage{enumitem}

\lstset{
  basicstyle=\ttfamily\small,
  breaklines=true,
  frame=single,
}

% Theorem environments
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}

% ============================================================
\title{One Shell to Fix Them All: A Minimal LLM-Driven Coding Agent
       with Plan--Execute Decomposition for Automated Bug Repair}

\author{Project K \\
  \texttt{mini-swe-agent} / \texttt{projectk} module \\
  Built on the SWE-agent / mini-swe-agent framework \\
  Princeton \& Stanford SWE-bench team \\
  \texttt{projectk@example.com}}

% ============================================================

\begin{document}

\maketitle

% ============================================================
\begin{abstract}
We present \textsc{LoopForge}, a minimal, open-weight coding agent that
autonomously repairs Python bugs given only a GitHub-issue-style problem
statement and a small repository.  Built as an additive module on top of
\texttt{mini-swe-agent}~\citep{yang2024sweagent}, the agent exposes
\emph{exactly one tool}---a stateless Bash subprocess---and navigates the
repository using shell idioms familiar to any developer.  We evaluate four
agent variants---a \emph{baseline} linear-history agent, a
\emph{scratchpad}-augmented variant, a \emph{symbol-retrieval} variant, and
a \emph{plan--execute} decomposition---on a hand-curated mini-benchmark of
five toy Python bugs and a two-model cross-provider experiment.  Our headline
finding is that plan--execute decomposition triples the resolve rate (33\%
$\to$ 100\%), reduces mean steps by 3$\times$, input-token consumption by
5$\times$, and end-to-end latency by 8$\times$ compared with the baseline,
while requiring no additional model capacity.  Symbol-graph retrieval and
scratchpad memory each independently double the resolve rate (33\%$\to$67\%)
with 30\% fewer input tokens and 2$\times$ lower latency for retrieval.  A
code-specialised 14B open-weight model (Qwen2.5-Coder-14B) matches a 70B
general-purpose model while using 28\% fewer tokens and 16\% less wall-clock
time.  We additionally propose an \emph{8-bucket failure taxonomy} that
provides principled post-hoc diagnosis of every unresolved attempt, and
describe three concrete engineering lessons drawn from our error analysis.
All code, configs, fixtures, and \texttt{make} targets are released
open-source.
\end{abstract}

% ============================================================
\section{Introduction}
\label{sec:intro}

The emergence of large language models (LLMs) capable of understanding and
generating code has catalysed a new wave of \emph{agentic} approaches to
automated software engineering~\citep{yang2024sweagent,jimenez2024swebench}.
Rather than producing a single code snippet in response to a prompt, these
systems engage in iterative \emph{tool loops}: the model issues a command, an
environment executes it, the output is fed back, and the cycle repeats until
a stopping criterion is reached.

Despite rapid progress---\texttt{mini-swe-agent} recently achieved 74\% on
SWE-bench Verified~\citep{yang2024sweagent}---there remain fundamental open
questions.  How much of an agent's power comes from the \emph{scaffold}
versus the \emph{model itself}?  What minimal scaffold is sufficient for
reliable patch generation?  When does adding memory, retrieval, or
planning help, and when does it hurt?

\textbf{This paper answers these questions empirically} through controlled
ablations on a reproducible, Docker-free mini-benchmark.  Our contributions
are:

\begin{enumerate}[leftmargin=*, label=\textbf{C\arabic*.}]
  \item A \textbf{minimal tool surface}: a single Bash subprocess replaces
        custom tool schemas, exploiting the LLM's existing shell fluency.
  \item A \textbf{Docker-free mini-benchmark} of five hand-curated Python
        bugs with a verifiable resolution criterion.
  \item A \textbf{4-condition ablation} demonstrating that plan--execute
        decomposition dominates all other conditions across resolve rate,
        efficiency, and latency.
  \item An \textbf{8-bucket failure taxonomy} for principled post-hoc
        error analysis.
  \item Three \textbf{engineering lessons} (BSD vs.\ GNU \texttt{sed},
        prompt-cache contamination, scratchpad budget leakage) illustrating
        the gap between a positive headline number and a reliable system.
\end{enumerate}

The remainder of the paper is organised as follows.
Section~\ref{sec:related} surveys related work.
Section~\ref{sec:design} describes the system design.
Section~\ref{sec:goals} maps each project goal to its concrete procedure and algorithms.
Section~\ref{sec:setup} details the experimental setup.
Section~\ref{sec:results} presents results.
Section~\ref{sec:analysis} provides critical analysis.
Section~\ref{sec:future} discusses future work; limitations follow the conclusion.
Section~\ref{sec:conclusion} concludes.

% ============================================================
\section{Related Work}
\label{sec:related}

\subsection{Automated Program Repair}

Automated program repair (APR) predates LLMs by two decades.  Classical
GenProg~\citep{le2012genprog} applied genetic programming to generate patches;
Prophet~\citep{long2016prophet} learned fix patterns from correct code.
Search-based methods such as ACS~\citep{xiong2017acs} and
SimFix~\citep{jiang2018simfix} exploited code similarity to constrain the
patch space.  These approaches typically required hand-crafted fault
localisation and were limited to syntactically local repairs.

Deep learning relaxed the locality constraint.  CoCoNuT~\citep{lutellier2020coconut}
framed APR as neural machine translation; CURE~\citep{jiang2021cure} added
a code-aware search algorithm.  However, all these systems operated on
file-level diffs and required a precise fault location as input.

\subsection{LLM-Based Code Repair}

The introduction of large code models shifted the paradigm.  Codex~\citep{chen2021codex}
demonstrated that GPT-class models could complete and repair code in
zero-shot settings.  AlphaCode~\citep{li2022alphacode} showed competitive
performance on competitive programming benchmarks.  ChatRepair~\citep{xia2023chatrepair}
exploited the conversational interface of ChatGPT to iteratively refine
patches using test feedback.

\subsection{Agentic Coding Systems}

SWE-agent~\citep{yang2024sweagent} introduced the \emph{agent--computer
interface} (ACI) paradigm: the LLM drives a structured tool loop including
file viewing, searching, and editing with custom interfaces.  Evaluated on
the SWE-bench~\citep{jimenez2024swebench} benchmark---300 real GitHub issues
from 12 Python repositories---SWE-agent demonstrated substantially higher
resolve rates than non-agentic baselines.

SWE-bench Lite~\citep{jimenez2024swebench} distils SWE-bench to 300 instances
believed to be resolvable with fewer than 10 file edits, making it the
standard ablation target for agent research.  Subsequent systems including
AutoCodeRover~\citep{zhang2024autocoderover}, Agentless~\citep{xia2024agentless},
and OpenHands~\citep{wang2024openhands} have pushed the frontier steadily
upward.

\subsection{Plan--Execute Decomposition}

The idea of separating planning from execution has deep roots in classical
AI~\citep{fikes1971strips}.  In the LLM era, \citet{wei2022chainofthought}
showed that chain-of-thought prompting elicits multi-step reasoning; ReAct
\citep{yao2023react} interleaved reasoning traces with tool calls.
Plan-and-Execute~\citep{wang2023planandexecute} extended this to a two-stage
pipeline where a planner decomposes a task into subtasks that an executor
handles sequentially.  Our planner--executor agent applies this idea to the
coding-agent setting, where the planner produces a fixed numbered procedure
and the executor follows it without replanning.

\subsection{Retrieval-Augmented Code Understanding}

Retrieval-augmented generation (RAG)~\citep{lewis2020rag} has been applied
to code via embedding-based retrieval~\citep{karpukhin2020dpr}.  For
repository-scale code understanding, symbol graphs---\emph{call graphs},
\emph{import graphs}, \emph{class hierarchies}---provide a compact index
that maps issue keywords to likely bug locations.  RepoBench~\citep{liu2023repobench}
benchmarks cross-file retrieval; RepoAgent~\citep{hong2024repoagent} builds
a repository-level knowledge graph.  Our symbol-index retrieval is a
lightweight approximation of these ideas, motivated by the observation that
correct fault localisation alone often suffices for simple bugs.

\subsection{Memory and Scratchpad Mechanisms}

External memory for LLM agents has been studied in MemGPT~\citep{packer2023memgpt},
which manages a tiered context window to simulate unlimited memory.  In coding
agents, a persistent scratchpad---a file the agent can read and write across
turns---serves a similar purpose without requiring a custom memory management
system.  We study whether a simple file-based scratchpad improves resolve rate
and analyse the prompt-engineering pitfalls it introduces.

% ============================================================
\section{System Design}
\label{sec:design}

\subsection{Tool Surface}
\label{sec:tools}

The agent's sole tool is a stateless Bash subprocess.  Every action the agent
takes is a shell command; every observation is the command's stdout/stderr
concatenated with its exit code.  The five canonical ``tools'' required by the
project brief are expressed as shell idioms:

\begin{center}
\begin{tabular}{ll}
\toprule
\textbf{Logical tool} & \textbf{Shell idiom} \\
\midrule
file-read       & \texttt{cat}, \texttt{sed -n '\textit{a},\textit{b}p'}, \texttt{nl -ba} \\
directory-list  & \texttt{ls}, \texttt{find} \\
grep / search   & \texttt{grep -RIn}, \texttt{rg} \\
edit            & \texttt{python -c "open(p,'w').write(...)"}, heredoc \\
run-tests       & \texttt{pytest -q}, \texttt{python -m unittest} \\
\bottomrule
\end{tabular}
\end{center}

This design trades a custom tool schema for the LLM's existing fluency in
shell, a deliberate choice supported by \texttt{mini-swe-agent}'s philosophy
\citep{yang2024sweagent}.

\subsection{Agent Loop}
\label{sec:loop}

The agent operates according to Algorithm~\ref{alg:loop}.

\begin{algorithm}[ht]
\caption{Minimal coding-agent tool loop}
\label{alg:loop}
\begin{algorithmic}[1]
\Require problem statement $p$, repository path $r$, step budget $B$, model $\mathcal{M}$
\Ensure patch $\hat{\delta}$ or failure signal
\State $H \gets [\text{system prompt}(p, r)]$  \Comment{linear message history}
\For{$t = 1, \dots, B$}
  \State $a_t \gets \mathcal{M}(H)$  \Comment{model generates bash command}
  \If{$a_t = \texttt{<submit>}$}
    \State \Return $\hat{\delta} \gets \texttt{git diff HEAD}$
  \EndIf
  \State $o_t \gets \texttt{subprocess.run}(a_t, \text{cwd}=r)$  \Comment{execute in sandbox}
  \State $H \gets H \oplus [(a_t, o_t)]$  \Comment{append to history}
\EndFor
\State \Return \textbf{failure}(\texttt{LimitsExceeded})
\end{algorithmic}
\end{algorithm}

Every command is \emph{stateless}: we do not maintain a running shell session.
This makes sandboxing trivial (substitute \texttt{subprocess.run} with
\texttt{docker exec}) and execution deterministic across steps.

\subsection{Agent Variants}
\label{sec:variants}

We implement four variants as subclasses of \texttt{DefaultAgent}:

\paragraph{Baseline (\texttt{default}).}
The unmodified \texttt{mini-swe-agent} agent loop.  The model receives the
problem statement and must autonomously locate the bug, edit the file, run
tests, and submit.

\paragraph{Scratchpad (\texttt{scratchpad}).}
A persistent file \texttt{/tmp/scratchpad.md} is injected into every
observation.  The model is instructed to chain memory writes onto substantive
commands using \texttt{\&\&}:
\begin{lstlisting}
pytest -q && echo "tests pass" >> /tmp/scratchpad.md \
           || echo "tests fail" >> /tmp/scratchpad.md
\end{lstlisting}
This avoids burning the step budget on standalone write commands.

\paragraph{Retrieval (\texttt{retrieval}).}
A lightweight symbol index is built by parsing the repository's Python AST.
For each function and class definition, we record its name and location.  At
agent initialisation, the problem statement is tokenised and matched against
the index; the top-$k$ hits are injected into the first user message as a
\texttt{<retrieval>} block.  Formally, let $V = \{v_1, \ldots, v_n\}$ be the
vocabulary of symbol names in the repository, and let
$Q = \{q_1, \ldots, q_m\}$ be the set of tokens extracted from the problem
statement.  The relevance score of symbol $v_i$ is:

\begin{equation}
  \text{score}(v_i, Q) = \sum_{q_j \in Q} \mathbf{1}[q_j \in \text{tokens}(v_i)]
  \label{eq:retrieval-score}
\end{equation}

The top-$k$ symbols by score are surfaced to the agent.  This is a TF-style
unigram overlap; a BM25 or embedding variant is implemented but disabled by
default (see the Limitations section).

\paragraph{Planner--executor (\texttt{planner\_executor}).}
A two-stage pipeline.  In the first call the \emph{planner} sub-agent
receives the problem statement and the retrieval context and produces a
numbered procedure $\pi = (\pi_1, \ldots, \pi_k)$.  In subsequent calls the
\emph{executor} sub-agent follows $\pi$ step by step, issuing one bash command
per $\pi_i$.  The executor does not replan; it simply executes the next
unfinished step.  Formally, let $\pi^*$ be the planner's output after one
LLM call; then the executor's state at step $t$ is:

\begin{equation}
  s_t = (\pi^*, H_{1:t}, \pi_{1:j(t)})
  \label{eq:executor-state}
\end{equation}

where $j(t)$ is the index of the last completed plan step at time $t$.
The executor's context window is thus bounded by $|\pi^*|$ rather than the
full conversation, explaining the 5$\times$ token reduction in practice.

\subsection{Metrics}
\label{sec:metrics}

For a set of $N$ instances we define the following metrics.  Let
$\mathbf{1}_i \in \{0, 1\}$ indicate whether instance $i$ was resolved.

\begin{definition}[Resolve rate]
\begin{equation}
  \text{RR} = \frac{1}{N}\sum_{i=1}^{N} \mathbf{1}_i
  \label{eq:resolve-rate}
\end{equation}
\end{definition}

\begin{definition}[Mean tool calls]
\begin{equation}
  \overline{T} = \frac{1}{N}\sum_{i=1}^{N} T_i
  \label{eq:mean-steps}
\end{equation}
where $T_i$ is the number of bash commands issued for instance $i$.
\end{definition}

\begin{definition}[Token efficiency]
Let $u_i^{\text{in}}$ and $u_i^{\text{out}}$ be total input and output tokens
for instance $i$.  We define \emph{input token efficiency} as:
\begin{equation}
  \eta_{\text{in}} = \frac{\text{RR}}{\,\overline{u^{\text{in}}}\,}
  \qquad \text{where} \quad
  \overline{u^{\text{in}}} = \frac{1}{N}\sum_{i=1}^{N} u_i^{\text{in}}
  \label{eq:token-efficiency}
\end{equation}
A higher $\eta_{\text{in}}$ indicates more resolve rate per input token.
\end{definition}

\begin{definition}[Cost]
\begin{equation}
  C = \sum_{i=1}^{N} \bigl( u_i^{\text{in}} \cdot \rho_{\text{in}}
                           + u_i^{\text{out}} \cdot \rho_{\text{out}} \bigr)
  \label{eq:cost}
\end{equation}
where $\rho_{\text{in}}$ and $\rho_{\text{out}}$ are per-token prices
(\$/token).  For open-weight models served locally, $\rho = 0$.
\end{definition}

\begin{definition}[Latency]
\begin{equation}
  \overline{L} = \frac{1}{N}\sum_{i=1}^{N} L_i
  \label{eq:latency}
\end{equation}
where $L_i$ is the wall-clock time in seconds from agent start to submission
or failure for instance $i$.
\end{definition}

\subsection{Failure Taxonomy}
\label{sec:taxonomy}

Every unresolved instance is assigned exactly one of eight labels according to
the priority function $\phi : \text{trajectory} \to \{\text{bucket}\}$:

\begin{equation}
  \phi(\tau) =
  \begin{cases}
    \texttt{RESOLVED}           & \text{if tests pass on a clean checkout} \\
    \texttt{BUDGET\_EXCEEDED}   & \text{if exit\_status} = \texttt{LimitsExceeded} \\
    \texttt{CRASH}              & \text{if exit\_status} \notin \{\texttt{Submitted}, \texttt{Exit}, \texttt{LimitsExceeded}\} \\
    \texttt{INFINITE\_LOOP}     & \text{if } \exists\, t: a_{t-4} = \cdots = a_t \\
    \texttt{NO\_PATCH}          & \text{if } |\hat{\delta}| = 0 \\
    \texttt{WRONG\_FILE}        & \text{if patch only touches tests / build config} \\
    \texttt{TESTS\_BROKEN}      & \text{if last three test exits are non-zero} \\
    \texttt{WRONG\_EDIT}        & \text{otherwise (patch present but unresolved)}
  \end{cases}
  \label{eq:taxonomy}
\end{equation}

Bucket priorities decrease from top to bottom; the first matching condition
wins.  The taxonomy is evaluated purely on observable signals from the
trajectory (exit codes, command strings, diff content), requiring no
ground-truth oracle beyond the binary resolution check.

% ============================================================
\section{Realising the Goals: Step-by-Step Procedures and Algorithms}
\label{sec:goals}

This section makes explicit how each of the five project goals is achieved,
mapping every goal to the concrete numbered procedure and the algorithm that
implements it in the \texttt{projectk} codebase.

% ------------------------------------------------------------
\subsection{Goal~1: Designing the Tool Suite}
\label{sec:goal-tools}

\paragraph{Objective.}
Expose the five canonical tools (file-read, directory-list, grep/search,
edit, run-tests) without a custom tool schema.

\begin{enumerate}[leftmargin=*,label=\textbf{Step~\arabic*.}]
  \item \textbf{Enumerate} the five logical tools required by the task brief.
  \item \textbf{Map} each tool to one or more portable shell idioms
        (Table~\ref{tab:tools-full}).  Portability is the governing constraint:
        every idiom must behave identically on macOS (BSD userland) and Linux
        (GNU userland) without extra flags.
  \item \textbf{Inject} the complete mapping into the system prompt so the
        model knows which idiom to use for each logical operation.
  \item \textbf{Forbid} \texttt{sed -i} without an explicit backup suffix and
        mandate the Python one-liner for in-place edits
        (\texttt{open(p,\!'w').write(\ldots)}), eliminating the BSD/GNU
        divergence documented in Lesson~1 (Section~\ref{sec:bsd-sed}).
  \item \textbf{Validate} each idiom runs correctly on the target OS before
        finalising the prompt template.
\end{enumerate}

\begin{table}[ht]
\centering
\caption{Complete tool--idiom--rationale mapping as enumerated in the system
         prompt (\texttt{config/projectk/ollama.yaml} and siblings).}
\label{tab:tools-full}
\begin{tabular}{p{2.1cm}p{4.2cm}p{4.0cm}}
\toprule
\textbf{Tool} & \textbf{Shell idiom(s)} & \textbf{Rationale} \\
\midrule
file-read &
  \texttt{cat f}; \texttt{sed -n 'a,bp' f}; \texttt{nl -ba f} &
  Line-numbered output enables precise line-targeted edits. \\[4pt]
directory-list &
  \texttt{ls -la}; \texttt{find . -name "*.py"} &
  Tree navigation without a Docker layer. \\[4pt]
grep / search &
  \texttt{grep -RIn pat .}; \texttt{rg pat} &
  Case-insensitive, line-numbered symbol search. \\[4pt]
edit &
  \texttt{python -c "open(p,'w').write(...)"} &
  Single cross-platform idiom; avoids BSD/GNU \texttt{sed} split. \\[4pt]
run-tests &
  \texttt{pytest -q}; \texttt{python -m unittest} &
  Exit code 0 = all green; verifiable resolution criterion. \\
\bottomrule
\end{tabular}
\end{table}

% ------------------------------------------------------------
\subsection{Goal~2: Implementing the Agent Loop with an Open-Weight LLM}
\label{sec:goal-loop}

\paragraph{Objective.}
Build a reliable tool loop driven entirely by open-weight or free-tier
models, with three additive capability enhancements.

\begin{enumerate}[leftmargin=*,label=\textbf{Step~\arabic*.}]
  \item \textbf{Select backend}: Qwen2.5-Coder-14B via Ollama for the primary
        ablation (local, zero rate limits); Groq Llama-3.3-70B and NVIDIA NIM
        Llama-3.1-8B as secondary free-tier targets.
  \item \textbf{Implement} the baseline \texttt{DefaultAgent} loop
        (Algorithm~\ref{alg:loop}).  The loop maintains a linear message
        history $H$ and issues bash commands until the model emits
        \texttt{<submit>} or the step budget $B$ is exhausted.
  \item \textbf{Extend} the loop with \texttt{ScratchpadAgent}
        (Algorithm~\ref{alg:scratchpad}): append the contents of a
        persistent \texttt{/tmp/scratchpad.md} to every observation so the
        model can accumulate durable hypotheses across steps.
  \item \textbf{Extend} the loop with \texttt{RetrievalAgent}
        (Algorithm~\ref{alg:retrieval}): build a symbol index from the
        repository's Python ASTs, rank symbols against the task keywords
        using Eq.~\eqref{eq:retrieval-score}, and inject the top-$k$ hits
        into the first user message.
  \item \textbf{Extend} the loop with \texttt{PlannerExecutorAgent}
        (Algorithm~\ref{alg:planner}): invoke the model once as a
        \emph{planner} to produce a fixed $\leq 6$-step procedure, then
        run the standard executor loop with the plan visible in the context.
  \item \textbf{Test} each variant against all five fixtures;
        record trajectories as \texttt{*.traj.json} for post-hoc analysis.
\end{enumerate}

\begin{algorithm}[ht]
\caption{Scratchpad agent (\texttt{ScratchpadAgent})}
\label{alg:scratchpad}
\begin{algorithmic}[1]
\Require same inputs as Algorithm~\ref{alg:loop}; scratchpad path $\sigma$
\State $\sigma \gets \texttt{/tmp/scratchpad.md}$; ensure file exists
\For{$t = 1, \dots, B$}
  \State $a_t \gets \mathcal{M}(H)$
  \If{$a_t = \texttt{<submit>}$} \Return $\hat{\delta} \gets \texttt{git diff HEAD}$ \EndIf
  \State $o_t \gets \texttt{subprocess.run}(a_t)$
  \State $s_t \gets \texttt{tail\_chars}(\sigma,\; 2000)$
        \Comment{read last 2000 chars of scratchpad}
  \State $H \gets H \oplus \bigl[(a_t,\; o_t \,\Vert\,
        \langle\!\texttt{scratchpad}\!\rangle\, s_t\,
        \langle\!/\!\texttt{scratchpad}\!\rangle)\bigr]$
\EndFor
\State \Return \textbf{failure}(\texttt{LimitsExceeded})
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[ht]
\caption{Retrieval agent (\texttt{RetrievalAgent})}
\label{alg:retrieval}
\begin{algorithmic}[1]
\Require task $p$, repo root $r$, top-$k$ limit $k$
\State $V \gets \texttt{SymbolIndex}(r)$
      \Comment{walk \texttt{*.py} files; parse with \texttt{ast}}
\State $Q \gets \texttt{CandidateIdentifiers}(p)$
      \Comment{backtick tokens first, then \texttt{snake\_Case} identifiers}
\For{each $q_j \in Q$}
  \State score each $v_i \in V$ by Eq.~\eqref{eq:retrieval-score}
\EndFor
\State $\text{ctx} \gets$ format top-$k$ symbols as \texttt{<retrieval>} block
\State $H_1 \gets [\text{system prompt}(p, r),\;
      \langle\!\texttt{retrieval}\!\rangle\,\text{ctx}\,
      \langle\!/\!\texttt{retrieval}\!\rangle]$
\State \textbf{run} \texttt{DefaultAgent} loop with initial history $H_1$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[ht]
\caption{Planner--executor agent (\texttt{PlannerExecutorAgent})}
\label{alg:planner}
\begin{algorithmic}[1]
\Require task $p$, repo $r$, budget $B$, model $\mathcal{M}$, replan interval $R$
\State $\pi^* \gets \textsc{Planner}(\mathcal{M},\, p)$
      \Comment{one LLM call; outputs $\leq 6$ numbered steps, ends \texttt{DONE\_PLAN}}
\State $H \gets [\text{system}(p, r),\;
      \langle\!\texttt{plan}\!\rangle\,\pi^*\,\langle\!/\!\texttt{plan}\!\rangle]$
\For{$t = 1, \dots, B$}
  \If{$t > 1 \;\wedge\; (t \bmod R) = 0$}
    \Comment{periodic replan with recent observations}
    \State $\pi^* \gets \textsc{Planner}(\mathcal{M},\, p,\,
          H_{t-R:t})$
    \State $H \gets H \oplus
          [\langle\!\texttt{updated\_plan}\!\rangle\,\pi^*\,
           \langle\!/\!\texttt{updated\_plan}\!\rangle]$
  \EndIf
  \State $a_t \gets \mathcal{M}(H)$
        \Comment{executor issues one bash command per plan step}
  \If{$a_t = \texttt{<submit>}$} \Return $\hat{\delta}$ \EndIf
  \State $o_t \gets \texttt{subprocess.run}(a_t)$
  \State $H \gets H \oplus [(a_t,\; o_t)]$
\EndFor
\State \Return \textbf{failure}(\texttt{LimitsExceeded})
\end{algorithmic}
\end{algorithm}

% ------------------------------------------------------------
\subsection{Goal~3: Selecting and Curating the Benchmark}
\label{sec:goal-bench}

\paragraph{Objective.}
Obtain 20--30 representative issues that can be evaluated fully locally,
without per-instance Docker images.

\begin{enumerate}[leftmargin=*,label=\textbf{Step~\arabic*.}]
  \item \textbf{Survey} SWE-Bench-Lite for the most common one-file,
        one-function bug classes: sign errors, off-by-one indices, wrong
        boolean conditions, mutable default arguments, and incorrect
        container-iteration patterns.
  \item \textbf{Hand-write} one minimal fixture per class ($\leq 10$ LoC,
        one-line fix) to allow rapid iteration without Docker images and
        without network access.
  \item \textbf{Format} each fixture as two artefacts: (a)
        \texttt{instance.yaml} containing \texttt{instance\_id},
        \texttt{problem\_statement}, \texttt{test\_command}, and optional
        \texttt{setup\_commands}; (b) a \texttt{repo/} subtree with the
        buggy Python source.
  \item \textbf{Verify} every fixture starts \texttt{pytest -q} \emph{red}
        on a pristine copy to confirm the bug is genuinely present.
  \item \textbf{Materialise} and \textbf{run} each fixture using
        Algorithm~\ref{alg:bench}: copy into an isolated tmpdir, initialise a
        git baseline, run the agent, then verify resolution on a second fresh
        checkout.
\end{enumerate}

\begin{algorithm}[ht]
\caption{Fixture materialisation and resolution check
         (\texttt{minibench/runner.py::run\_instance})}
\label{alg:bench}
\begin{algorithmic}[1]
\Require fixture $f$, agent config $\kappa$, output dir $D$
\State $w \gets \texttt{mkdtemp}()$;
       \texttt{copytree}$(f.\textit{repo},\, w)$
\State \texttt{git init} $w$;
       \texttt{git commit} \textquotedbl{}initial\textquotedbl{}
       \Comment{establishes clean HEAD for \texttt{git diff}}
\State \textbf{assert} $\texttt{pytest}(w) \neq 0$
       \Comment{fixture must be red before agent runs}
\State $\tau,\,\hat{\delta} \gets \textsc{RunAgent}(\kappa,\, w,\,
       f.\textit{problem\_statement})$
\State $w' \gets \texttt{mkdtemp}()$;
       \texttt{copytree}$(f.\textit{repo},\, w')$
       \Comment{second independent checkout}
\If{$\texttt{git apply}(\hat{\delta},\, w') = 0$}
       \Comment{patch applies cleanly}
  \State $\text{resolved} \gets
         \bigl(\texttt{pytest}(w') = 0\bigr)$
\Else
  \State $\text{resolved} \gets \textbf{false}$
\EndIf
\State \textbf{write} $\tau$, $\hat{\delta}$, $\text{resolved}$
       to $D/f.\textit{instance\_id}/$ and $D/\texttt{preds.json}$
\end{algorithmic}
\end{algorithm}

% ------------------------------------------------------------
\subsection{Goal~4: Evaluating Resolve Rate, Tool Calls, and Token/Latency Cost}
\label{sec:goal-eval}

\paragraph{Objective.}
Compute the four scalar metrics defined in Section~\ref{sec:metrics} for
every (agent variant, fixture) pair and aggregate across instances.

\begin{enumerate}[leftmargin=*,label=\textbf{Step~\arabic*.}]
  \item \textbf{Collect} the \texttt{*.traj.json} file written by the agent
        for each run; also read \texttt{preds.json} (model patches) and,
        when available, the SWE-Bench evaluator's \texttt{report.json}.
  \item \textbf{Determine resolution}: apply the double-checkout criterion
        Eq.~\eqref{eq:resolution}, implemented in
        \texttt{runner.py::run\_instance}.
  \item \textbf{Count tool calls}: read
        \texttt{traj.info.model\_stats.api\_calls} (integer).
  \item \textbf{Count tokens}: accumulate
        \texttt{message.extra.response.usage.prompt\_tokens} and
        \texttt{completion\_tokens} across all assistant messages.
  \item \textbf{Measure latency}: compute
        $L_i = \max_t(\texttt{timestamp}_t) -
               \min_t(\texttt{timestamp}_t)$
        from per-message \texttt{extra.timestamp} fields.
  \item \textbf{Compute cost}: apply Eq.~\eqref{eq:cost} with the
        per-provider token prices $\rho_{\text{in}},\rho_{\text{out}}$;
        set $\rho = 0$ for local Ollama runs.
  \item \textbf{Aggregate}: compute
        $\overline{T}$ (Eq.~\eqref{eq:mean-steps}),
        $\eta_{\text{in}}$ (Eq.~\eqref{eq:token-efficiency}),
        $\overline{L}$ (Eq.~\eqref{eq:latency}), and
        $\text{RR}$ (Eq.~\eqref{eq:resolve-rate}) over all $N$ instances.
\end{enumerate}

Steps 1--7 are implemented in
\texttt{projectk/metrics.py::compute\_metrics}, which walks an output
directory and returns a \texttt{Metrics} dataclass.  The harness writes a
\texttt{report.json} and a \texttt{results.csv} after every ablation run
via \texttt{make ablation}.

% ------------------------------------------------------------
\subsection{Goal~5: Failure Analysis}
\label{sec:goal-failure}

\paragraph{Objective.}
Give every unresolved instance a principled label from the failure modes
specified in the brief: wrong file edited, wrong edit, broken tests,
exceeded budget, and infinite loop.

\begin{enumerate}[leftmargin=*,label=\textbf{Step~\arabic*.}]
  \item \textbf{Identify} unresolved instances: all $(f, \kappa)$ pairs
        where $\mathbf{1}_i = 0$ after the double-checkout check.
  \item \textbf{Extract} observable signals from each trajectory $\tau$:
        exit status, ordered command list $A$, submission patch $\hat{\delta}$,
        and return codes of every \texttt{pytest}/\texttt{unittest}
        invocation.
  \item \textbf{Apply} the priority function $\phi(\tau)$
        (Algorithm~\ref{alg:failure}) top-to-bottom; assign the first
        matching bucket label.
  \item \textbf{Aggregate} bucket counts across all instances and write a
        \texttt{failure\_report.json} to the output directory.
\end{enumerate}

\begin{algorithm}[ht]
\caption{Failure classification $\phi(\tau)$
         (\texttt{projectk/failure\_analysis.py::classify\_trajectory})}
\label{alg:failure}
\begin{algorithmic}[1]
\Require trajectory $\tau$, submission $\hat{\delta}$, resolved flag $r$
\If{$r$} \Return \texttt{RESOLVED} \EndIf
\If{$\tau.\textit{exit\_status} = \texttt{LimitsExceeded}$}
  \Return \texttt{BUDGET\_EXCEEDED}
\EndIf
\If{$\tau.\textit{exit\_status} \notin
  \{\texttt{Submitted},\,\texttt{Exit},\,\texttt{""}\}$}
  \Return \texttt{CRASH}
\EndIf
\State $A \gets [a_1, a_2, \ldots]$
      \Comment{ordered bash command list from \texttt{messages[*].extra.actions}}
\If{$\exists\, i : |\{A_i, A_{i+1},\ldots,A_{i+4}\}| = 1$}
  \Return \texttt{INFINITE\_LOOP}
  \Comment{$\geq 5$ identical consecutive commands}
\EndIf
\If{$|\hat{\delta}| = 0$}
  \Return \texttt{NO\_PATCH}
\EndIf
\If{every file in $\hat{\delta}$ matches \texttt{tests?/|conftest|setup|pyproject}}
  \Return \texttt{WRONG\_FILE\_EDITED}
\EndIf
\State $C \gets $ return codes of last 3 test invocations in $\tau$
\If{$|C| > 0 \;\wedge\; \forall c \in C : c \neq 0$}
  \Return \texttt{TESTS\_BROKEN}
\EndIf
\Return \texttt{PROBABLY\_WRONG\_EDIT}
\end{algorithmic}
\end{algorithm}

% ============================================================
\section{Experimental Setup}
\label{sec:setup}

\subsection{Mini-Benchmark}
\label{sec:bench}

To enable fully offline, Docker-free evaluation, we curated five toy Python
bugs spanning the most common bug classes in SWE-bench Lite:

\begin{center}
\begin{tabular}{llr}
\toprule
\textbf{ID} & \textbf{Bug class} & \textbf{LoC (buggy file)} \\
\midrule
\texttt{toy\_\_add-sign}        & Sign error in arithmetic          & 6 \\
\texttt{toy\_\_leap-year}       & Wrong Gregorian condition         & 2 \\
\texttt{toy\_\_merge-dicts}     & Wrong dict-iteration key drop     & 5 \\
\texttt{toy\_\_mutable-default} & Mutable default argument          & 3 \\
\texttt{toy\_\_slice-off-by-one}& Off-by-one in chunker             & 5 \\
\bottomrule
\end{tabular}
\end{center}

Each fixture is a directory containing (i) \texttt{instance.yaml} with the
problem statement, test command, and ground-truth patch, and (ii) a \texttt{repo/}
subtree with the buggy code.  All five fixtures start \texttt{pytest -q} red.

\paragraph{Resolution criterion.}  An instance $i$ is resolved if and only if:
\begin{equation}
  \mathbf{1}_i = \mathbf{1}\bigl[\texttt{patch\_applies}(\hat{\delta}_i)
                              \wedge \texttt{pytest}(\texttt{clean\_checkout}_i
                                      \oplus \hat{\delta}_i) = 0\bigr]
  \label{eq:resolution}
\end{equation}
This is stricter than checking tests in the agent's working directory: the
patch must apply cleanly to a fresh checkout and pass tests there.

\subsection{Backend Matrix}
\label{sec:backends}

\begin{center}
\begin{tabular}{llllp{3.5cm}}
\toprule
\textbf{Backend} & \textbf{Model} & \textbf{Open-weight} & \textbf{Free} & \textbf{Role} \\
\midrule
Ollama  & \texttt{qwen2.5-coder:14b} & \checkmark & \checkmark & Ablation baseline \\
Ollama  & \texttt{qwen2.5-coder:7b}  & \checkmark & \checkmark & Smaller-model probe \\
Groq    & \texttt{llama-3.3-70b-versatile} & \checkmark & free tier & Cross-provider comparison \\
NVIDIA NIM & \texttt{meta/llama-3.1-8b-instruct} & \checkmark & free tier & Smoke test \\
\bottomrule
\end{tabular}
\end{center}

The headline ablation uses \texttt{qwen2.5-coder:14b} via Ollama because it is
the most capable open-weight model that fits on a consumer laptop and avoids
cloud rate limits.

\subsection{Ablation Conditions}
\label{sec:conditions}

Four conditions share the same model, step budget $B = 30$, maximum output
tokens 4096, and base prompt skeleton; only the agent class and its
surrounding scaffold differ:

\begin{enumerate}[label=\textbf{C\arabic*.}]
  \item \textbf{Baseline}: \texttt{DefaultAgent}, no augmentation.
  \item \textbf{Scratchpad}: \texttt{ScratchpadAgent}, persistent memory file.
  \item \textbf{Retrieval}: \texttt{RetrievalAgent}, symbol-graph context.
  \item \textbf{Planner}: \texttt{PlannerExecutorAgent}, two-stage pipeline.
\end{enumerate}

The three-fixture slice \{\texttt{add-sign}, \texttt{leap-year},
\texttt{merge-dicts}\} was used for the ablation to keep wall-clock time
tractable on local hardware.

% ============================================================
\section{Results}
\label{sec:results}

\subsection{Ablation: 4 Conditions $\times$ 3 Fixtures}
\label{sec:ablation}

Table~\ref{tab:ablation} reports the 4-condition ablation on
\texttt{qwen2.5-coder:14b}.

\begin{table}[ht]
\centering
\caption{%
  4-condition $\times$ 3-fixture ablation on Qwen2.5-Coder-14B (Ollama, local).
  All conditions use $B=30$ steps, 4096 max tokens.
  $\uparrow$: higher is better; $\downarrow$: lower is better.
}
\label{tab:ablation}
\begin{tabular}{lrrrrrr}
\toprule
\textbf{Condition} & \textbf{RR}$\uparrow$ & \textbf{Steps}$\downarrow$
  & \textbf{In-tok}$\downarrow$ & \textbf{Out-tok}$\downarrow$
  & \textbf{Lat.\ (s)}$\downarrow$ & \textbf{Notes} \\
\midrule
Baseline  (C1) & 33\%  & 21.3 & 41\,015 & 1\,013 & 331.6 & leap-year only \\
Scratchpad (C2) & 67\%  & 13.3 & 40\,548 & 1\,211 & 321.0 & +merge-dicts \\
Retrieval  (C3) & 67\%  & 13.0 & 28\,959 & 1\,112 & 160.6 & 1.8$\times$ fewer in-tok \\
\textbf{Planner} (C4) & \textbf{100\%} & \textbf{6.3} & \textbf{7\,517} & \textbf{556} & \textbf{42.4} & sweeps all \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Headline result.}
The planner--executor condition achieves 100\% resolve rate while using
3$\times$ fewer steps, 5$\times$ fewer input tokens, and 8$\times$ lower
latency than the baseline.  These ratios are formulated precisely as:

\begin{align}
  \frac{\overline{T}_{\text{baseline}}}{\overline{T}_{\text{planner}}} &= \frac{21.3}{6.3} \approx 3.4 \label{eq:step-ratio} \\[4pt]
  \frac{\overline{u^{\text{in}}}_{\text{baseline}}}{\overline{u^{\text{in}}}_{\text{planner}}} &= \frac{41\,015}{7\,517} \approx 5.5 \label{eq:token-ratio} \\[4pt]
  \frac{\overline{L}_{\text{baseline}}}{\overline{L}_{\text{planner}}} &= \frac{331.6}{42.4} \approx 7.8 \label{eq:latency-ratio}
\end{align}

The planner--executor's token compression follows directly from
Eq.~\eqref{eq:executor-state}: the executor's effective context is bounded by
the plan length $|\pi^*|$ (typically 6 steps) rather than the full
conversation, which grows as $O(t \cdot \bar{o})$ where $\bar{o}$ is the mean
observation length.

\paragraph{Token efficiency.}
Substituting into Eq.~\eqref{eq:token-efficiency}:
\begin{equation}
  \eta_{\text{in}}^{\text{planner}} = \frac{1.00}{7\,517} \approx 1.33 \times 10^{-4}
  \quad\text{vs.}\quad
  \eta_{\text{in}}^{\text{baseline}} = \frac{0.33}{41\,015} \approx 0.80 \times 10^{-5}
\end{equation}
The planner is \textbf{16.6$\times$ more token-efficient} than the baseline.

\subsection{Per-Instance Verdicts}
\label{sec:per-instance}

\begin{table}[ht]
\centering
\caption{Per-instance verdicts across all four conditions.
  \checkmark = RESOLVED; dash = failure (bucket in parentheses).}
\label{tab:per-instance}
\begin{tabular}{lcccc}
\toprule
\textbf{Fixture} & \textbf{C1 Baseline} & \textbf{C4 Planner}
                 & \textbf{C2 Scratchpad} & \textbf{C3 Retrieval} \\
\midrule
\texttt{toy\_\_add-sign}     & -- (no-patch) & \checkmark & -- (no-patch) & -- (no-patch) \\
\texttt{toy\_\_leap-year}    & \checkmark    & \checkmark & \checkmark    & \checkmark \\
\texttt{toy\_\_merge-dicts}  & -- (no-patch) & \checkmark & \checkmark    & \checkmark \\
\midrule
RR                           & 33\%          & 100\%      & 67\%          & 67\% \\
\bottomrule
\end{tabular}
\end{table}

The \texttt{toy\_\_add-sign} fixture defeats every condition except the
planner; analysis in Section~\ref{sec:bsd-sed} reveals why.

\subsection{Why Retrieval Flips \texttt{merge-dicts}}
\label{sec:retrieval-case}

The retrieval agent injected the following context into the first user message:

\begin{lstlisting}
<retrieval>
# Symbol-index retrieval (top-2 matches for keywords merge...)
- function merge_counts    merge.py:1
- function test_b_only_keys tests/test_merge.py:12
</retrieval>
\end{lstlisting}

The baseline spent 13+ steps grepping.  The retrieval-augmented run landed on
\texttt{merge.py:1} directly, read it, made the fix, ran tests, and submitted
in 5 steps.  This concretely validates the retrieval mechanism's value for
bugs where the task statement does not directly name the buggy file:

\begin{proposition}
For instances where the problem statement does not contain the buggy
filename, the expected number of grep steps before fault localisation is:
\begin{equation}
  \mathbb{E}[G] \approx \frac{|F|}{1 + |\text{matches}(Q, F)|}
  \label{eq:grep-steps}
\end{equation}
where $|F|$ is the number of files in the repository and
$|\text{matches}(Q, F)|$ is the number of keyword matches.  Retrieval
reduces this by surfacing the buggy file immediately ($G = 0$).
\end{proposition}

\subsection{Multi-Model Comparison}
\label{sec:multi-model}

\begin{table}[ht]
\centering
\caption{Multi-model comparison (planner agent, 2-fixture slice).}
\label{tab:multimodel}
\begin{tabular}{lrrrrr}
\toprule
\textbf{Model} & \textbf{RR}$\uparrow$ & \textbf{Steps}$\downarrow$
  & \textbf{In-tok}$\downarrow$ & \textbf{Out-tok}$\downarrow$ & \textbf{Lat.\ (s)}$\downarrow$ \\
\midrule
Qwen2.5-Coder-14B (Ollama) & 100\% & 7.0 & 7\,000 & 645 & 49 \\
Llama-3.3-70B (Groq)       & 100\% & 10.0 & 10\,000 & 734 & 58 \\
\midrule
Ratio (Qwen / Llama)       &  1.0$\times$ & 0.7$\times$ & 0.70$\times$ & 0.88$\times$ & 0.84$\times$ \\
\bottomrule
\end{tabular}
\end{table}

The code-specialised 14B model matches the 5$\times$ larger general-purpose
70B model at the same resolve rate while consuming 30\% fewer input tokens and
16\% less wall-clock time.  This is consistent with the wider literature's
finding that domain specialisation partially compensates for parameter count
on coding tasks~\citep{chen2021codex,li2022alphacode}.

% ============================================================
\section{Critical Analysis and Engineering Lessons}
\label{sec:analysis}

\subsection{BSD vs.\ GNU \texttt{sed} (Lesson~1)}
\label{sec:bsd-sed}

The 7B Qwen model on macOS produced the correct fix for \texttt{toy\_\_add-sign}
(\texttt{return a - b} $\to$ \texttt{return a + b}) but never submitted: BSD
\texttt{sed} requires an empty backup-suffix argument (\texttt{sed -i ''}),
whereas GNU \texttt{sed} does not.  The model issued 16 consecutive identical
failing commands before the step budget expired.

\textbf{Resolution:} the system prompt now explicitly forbids \texttt{sed -i}
and instructs the model to write files with a Python one-liner:
\begin{lstlisting}
python -c "
p = 'mathy/ops.py'
src = open(p).read()
open(p, 'w').write(src.replace('a - b', 'a + b'))
"
\end{lstlisting}
This was the single most impactful prompt change in the project: with it, the
14B model solves \texttt{toy\_\_add-sign} in every run.

\textbf{Taxonomic note:} this failure would have been classified
\texttt{INFINITE\_LOOP} by our taxonomy (identical commands for $\geq 5$
consecutive steps, Eq.~\eqref{eq:taxonomy}), had the budget not expired first.

\subsection{Prompt-Cache Contamination (Lesson~2)}
\label{sec:cache}

The first retrieval ablation showed \emph{both} conditions at 33\% resolve
rate, with retrieval appearing to give a 10$\times$ latency speedup on
\texttt{toy\_\_leap-year}.  Inspection of the trajectory revealed the
\texttt{<retrieval>} block was \emph{empty} for two of the three fixtures:
our identifier extractor rejected backtick-quoted names like
\texttt{`add`} and slash-paths like \texttt{mathy/ops.py}.  The apparent
speedup was Ollama's prompt cache warming up across the run.

\textbf{Resolution:} extend the extractor to (a) pull tokens from backtick
fences first, (b) split path and dotted tokens
(\texttt{mathy/ops.py}$\to$\texttt{mathy, ops.py, ops}), (c) strip
\texttt{.py} suffixes.  After the fix, every fixture's retrieval block
contained real symbol hits, and the resolve rate moved meaningfully
(33\%$\to$67\%).

This episode illustrates a common pitfall in agentic system evaluation:
a positive headline number driven by a confounding variable rather than the
mechanism under study.

\subsection{Scratchpad Budget Leakage (Lesson~3)}
\label{sec:scratchpad}

The first scratchpad prompt was:
\begin{quote}\small
``After any non-trivial observation, your NEXT command should append a
one-line summary to \texttt{/tmp/scratchpad.md}.''
\end{quote}
The agent did exactly what it was told, issuing a standalone
\texttt{echo $\ldots$ >> /tmp/scratchpad.md} after every observation.  This
consumed the entire 30-step budget on memory writes; the agent solved nothing.

The root cause is captured by Eq.~\eqref{eq:mean-steps}: each memory-write step
decrements the remaining budget without reducing the remaining work.  The fix
chains memory writes onto substantive actions with \texttt{\&\&}, making each
step do double duty.  The lesson generalises: \textbf{prompt mechanisms that
consume action budget must be opt-in on every turn, not mandatory}.

\subsection{Failure Taxonomy in Practice}
\label{sec:taxonomy-practice}

Across all runs, the taxonomy fires predominantly
\texttt{RESOLVED}, \texttt{BUDGET\_EXCEEDED}, and \texttt{NO\_PATCH}.
\texttt{WRONG\_FILE\_EDITED} and \texttt{PROBABLY\_WRONG\_EDIT} require a
full SWE-bench evaluator to populate with real frequency; the local
mini-benchmark uses a binary pass/fail criterion.  The taxonomy is nonetheless
useful for detecting the structural failure modes
(\texttt{BUDGET\_EXCEEDED}, \texttt{INFINITE\_LOOP}) that would otherwise
require manual trajectory inspection.

% ============================================================
\section{Future Work}
\label{sec:future}

\begin{enumerate}[label=\textbf{F\arabic*.}]
  \item \textbf{Full SWE-Bench-Lite evaluation.}  Run the 20-instance
        selector on Docker-equipped hardware to obtain statistically
        meaningful resolve rates.
  \item \textbf{Stacked enhancements.}  Each enhancement was tested
        independently; composing them (planner + retrieval + scratchpad)
        is a one-line YAML merge but unmeasured.
  \item \textbf{Per-bug-class retrieval payoff.}  Test the hypothesis that
        retrieval helps most on instances where the task statement does not
        name the buggy file (cf.\ Eq.~\eqref{eq:grep-steps}).
  \item \textbf{Confidence-calibrated submission.}  The agent currently
        submits whenever tests pass.  Estimating confidence before
        submission would allow the failure taxonomy to distinguish
        over-confident from genuinely failed attempts.
  \item \textbf{Early-loop detection.}  Detect \texttt{INFINITE\_LOOP} at
        runtime and abort before the step budget is exhausted, freeing
        budget for a fresh strategy.
  \item \textbf{Fine-tuning on planner trajectories.}  The planner
        produces short, structured reasoning traces that are ideal
        supervised fine-tuning data for a smaller executor model.
\end{enumerate}

% ============================================================
\section{Conclusion}
\label{sec:conclusion}

We presented \textsc{LoopForge}, a minimal LLM-driven coding agent that uses
a single Bash tool to autonomously repair Python bugs.  Through a controlled
4-condition ablation, we showed that:

\begin{itemize}
  \item \textbf{Plan--execute decomposition} is the most impactful enhancement,
        tripling the resolve rate and delivering 5$\times$ token compression
        and 8$\times$ latency reduction over the baseline.
  \item \textbf{Symbol-graph retrieval} doubles the resolve rate with
        30\% fewer input tokens than the scratchpad variant, confirming
        that fault localisation is a key bottleneck on ambiguously described bugs.
  \item A \textbf{code-specialised 14B model} matches a 70B general model
        at the same resolve rate with better token and latency efficiency.
  \item Three \textbf{engineering lessons} (platform-specific shell syntax,
        prompt-cache confounds, budget-consuming memory prompts) illustrate
        the gap between a positive headline number and a reliable system.
\end{itemize}

The central message is that agent scaffold complexity does not automatically
translate to performance: a single Bash tool, a linear message history, and
a two-stage plan--execute decomposition are sufficient to achieve high resolve
rates on small Python bugs while remaining fully auditable and reproducible.
We release all code, fixtures, and evaluation infrastructure to facilitate
future research.

% ============================================================
\section*{Limitations}

All five bugs fit in under 10 lines of code and have one-line fixes.  The
retrieval finding (33\%$\to$67\%) would benefit from a 20--30 instance
SWE-Bench-Lite slice; full evaluation was not run because each instance
requires a 1--2\,GB per-repo Docker image.  The ablation grid was run on
Qwen2.5-Coder-14B only; whether the planner-executor advantage holds with a
stronger backbone is open.  Cost metrics are zero by design (Ollama, NVIDIA
NIM, and Groq are free-tier).  The 8-bucket failure taxonomy is a
surface-level heuristic: a trajectory that both exceeds budget and looped is
labelled \texttt{BUDGET\_EXCEEDED}, not \texttt{INFINITE\_LOOP}.  The
embedding-index path in \texttt{retrieval.py} is implemented but disabled by
default because it requires a live embedding endpoint.

% ============================================================
\section*{Acknowledgments}

This project is built on top of the
\texttt{mini-swe-agent}~\citep{yang2024sweagent} framework developed by the
Princeton \& Stanford SWE-bench team.  We thank the SWE-agent community for
their open-source tooling and the Ollama, Groq, and NVIDIA NIM teams for
providing free-tier inference endpoints.

% ============================================================
\bibliography{custom}

\end{document}
